\textbf{{\Large {Recurrent relation for branching coefficients of affine Lie algebras}}}

### 2009-12-08 AffineLieAlgebras/InformalNotes/article.tex
Added some text on the construction of $\mathfrak{a}_{\bot}$

If $\mathfrak{a}$ is a maximal regular subalgebra of $\mathfrak{g}$ then rank of $\mathfrak{a}$ is equal to the rank of $\mathfrak{g}$ and it is clear that $\Delta^{+}_{\bot}$ is empty.
\begin{comment}
  Analyze all the regular embeddings and show when there exist orthogonal roots 
\end{comment}
Non-maximal regular embedding of $\mathfrak{a}$ into $\mathfrak{g}$ can be obtained through the chain of maximal embeddings $\mathfrak{a}\subset \mathfrak{p}_1\subset \mathfrak{p}_2\subset\dots \subset \mathfrak{g}$. Also the maximal regular embeddings are constructed by the exclusion of one or two roots from the extended Dynkin diagram of the algebra. Since this process can give us non-connected Dynkin diagrams we can see which roots are orthogonal to the root space of non-maximal regular subalgebra $\mathfrak{a}$.

Consider for instance regular embedding of $A_1\subset B_2$ ($su(2)\subset so(5)$).

@@

In the case of special embeddings the set $\Delta^{+}_{\bot}$ can be empty as for the special embedding of $A_1\subset A_2$ with the embedding index equal to 4, or non-empty for example for the embedding $su(p)\subset su(p)\oplus su(q)\subset $
\begin{comment}
  Add here the example
\end{comment}

@@

Let's describe the reduction of fundamental representation of $B_2$ with the highest weight (in fundamental weight basis) equal to $(1,0)$.
\begin{figure}[ph]
  \noindent\centering{
    \includegraphics[width=90mm]{B2_A1.pdf}
  }
  \caption{Regular embedding of $A_1$ into $B_2$}
  \label{fig:B2_A1}
\end{figure}
On the Figure \ref{fig:B2_A1} we have also shown the set of points $\omega(\mu+\rho),\; \omega\in W$ of fundamental representation of $B_2$ with the corresponding determinants of Weyl reflections $\epsilon(\omega)$. 
Now we have to factorise the Weyl group $W$ by $W_{\bot}=\left\{\omega_1\right\}$. We get the following set of anomalous points $\omega(\mu+\rho)-\rho,\; \omega\in W_{\bot}\backslash W$:
\begin{figure}[ph]
  \noindent\centering{
    \includegraphics[width=90mm]{B2_A1_2.pdf}
  }
  \caption{Anomalous points and the corresponding $\mathfrak{a}_{\bot}=A_1$-modules}
  \label{fig:B2_A1_2}
\end{figure}
We have also depicted the corresponding $\mathfrak{a}_{\bot}=A_1$-modules $L^{\pi_{\mathfrak{a}_{\bot}}(\omega(\mu+\rho))-\rho_{\mathfrak{a}_{\bot}}}_{\mathfrak{a}_{\bot}}$.
Then we project these points and dimensions of modules onto the root space of subalgebra $\mathfrak{a}=A_1$ and get the following anomalous points in fundamental weights basis with corresponding multiplicities:
\begin{equation}
  \label{eq:25}
  (1,2),\; (0,-3),\; (-4,3),\; (-5,-2)
\end{equation}
For the function $s(\gamma)$ and the set $\Gamma$ from the definition (\ref{phi-d},\ref{fan-defined}) we have
\begin{equation}
  \label{eq:22}
  (0,1),\; (1,2),\; (2,-1)
\end{equation}
Here the second component denotes the value of $s(\gamma)$.
\begin{comment}
  It seems that here $\Gamma$ should include only two points. Is it true?

  Think of the pictures more.
\end{comment}
Anomalous branching coefficient $k^{(1,0)}_{1}=2$, then for anomalous branching coefficient $k^{(1,0)}_{0}$ the formula (\ref{recurrent relation}) gives us
\begin{equation}
  \label{eq:23}
  k^{(1,0)}_{0}=-1\cdot k^{(1,0)}_2 +2\cdot k^{(1,0}_1 - 3\cdot \delta_{0,0} = 1
\end{equation}
So we have computed the branching coefficients.
\subsubsection{Embedding of $B_2$ into $B_4$}
\label{sec:someth-high-dimens}
Consider the the regular embedding of the subalgebra $B_2$ into the algebra $B_4$.
We calculate the branching coefficients for the fundamental representation of $B_4$.
The corresponding Dynkin diagrams are in the Figure \ref{fig:dynkin}.
\begin{figure}[ph]
  \centering
  \includegraphics[width=60mm]{B4_B2_2A1.pdf}
  \caption{Dynkin diagrams}
  \label{fig:dynkin}
\end{figure}

>>>>

If $\mathfrak{a}$ is a maximal regular subalgebra of $\mathfrak{g}$ then rank of $\mathfrak{a}$ is equal to the rank of $\mathfrak{g}$ and it is clear that $\Delta^{+}_{\bot}$ is empty.
On the other hand 
non-maximal regular embedding of $\mathfrak{a}$ into $\mathfrak{g}$ can be obtained through the chain of maximal embeddings $\mathfrak{a}\subset \mathfrak{p}_1\subset \mathfrak{p}_2\subset\dots \subset \mathfrak{g}$. Also the maximal regular embeddings are constructed by the exclusion of one or two roots from the extended Dynkin diagram of the algebra. Since this process can give us non-connected Dynkin diagrams we can see which roots are orthogonal to the root space of non-maximal regular subalgebra $\mathfrak{a}$.

Consider for instance regular embedding of $A_1\subset B_2$.

@@

Note that when we study the regular embedding obtained by the dropping of the nodes of the extended Dynkin diagram of the algebra $\mathfrak{g}$ and the subalgebra $\mathfrak{a}$ is one of the connected components, the subalgebra $\mathfrak{a}_{\bot}$ may be broader than the algebra generated by the other connected components. See for example the figure \ref{fig:dynkin} for the embedding of $B_2\subset B_4$. 
In this case the extended Dynkin diagram of $B_4$ is broken by the dropping of simple root $\alpha_2=e_2-e_3$ into the diagrams of the subalgebra of interest $\mathfrak{a}=B_2$ and two diagrams of the orthogonal $A_1$ subalgebras. But the subalgebra $\mathfrak{a}_{\bot}$ is equal not to $A_1\oplus A_1$ but to $B_2$. 
This fact is due to the existence of positive root $e_2=s_{\alpha_2}\cdot s_{\alpha_3}\cdot\alpha_4$ in the root system of $B_4$.

In the language of Weyl groups the task of construction of the root system of $\mathfrak{a}_{\bot}$ is formulated as follows. Given a Weyl group $W$ and a subgroup $W_{\mathfrak{a}}$ find the subgroup $W_{\mathfrak{a}_{\bot}}\subset W$ such that $\omega\, W_{\mathfrak{a}}\,\omega^{-1}=W_{\mathfrak{a}}; \;\text{for}\; \omega\in W_{\mathfrak{a}_{\bot}}$, so we need to find the normalizer of $W_{\mathfrak{a}}$ in $W$. This task can be carried out using the computational methods of finite group theory \cite{stembridge2001computational} but we are unaware of any publication with the exhaustive tables of such a subgroups.

In the case of special embeddings the set $\Delta^{+}_{\bot}$ can be empty as for the special embedding of $A_1\subset A_2$ with the embedding index equal to 4, or non-empty for example for the embedding $A_1\subset A_2\subset A_3$ which is depicted at the Figure \ref{fig:a1a2a3}.
\begin{figure}[ph]
  \centering
  \includegraphics[width=120mm]{special_A1_A2_A3.pdf}
  \caption{Special embedding $A_1\subset A_2\subset A_3$.\\ Grey vectors are the roots of $A_2$, thick black - of $\mathfrak{a}=A_1$, \\dashed black are the orthogonal roots of $A_1$ which is contained in $\mathfrak{a}_{\bot}$ }
  \label{fig:a1a2a3}
\end{figure}

The construction of the subalgebra $\mathfrak{a}_{\bot}$ is not obvious from the Dynkin diagrams of $\mathfrak{g}$ and $\mathfrak{a}$ but for the needs of calculation of the branching coefficients $b^{(\mu)}_{\nu}$ it is sufficient to select the positive roots $\Delta^{+}_{\bot}=\left\{\alpha\in \Delta^{+}:\alpha\bot \Delta_{\mathfrak{a}}\right\}$ in the set of positive roots of $\mathfrak{g}$.

@@

Let's describe the reduction of fundamental representation of $B_2$ with the highest weight (in fundamental weight basis) equal to $(1,0)$.
\begin{figure}[ph]
  \noindent\centering{
    \includegraphics[width=90mm]{B2_A1.pdf}
  }
  \caption{Regular embedding of $A_1$ into $B_2$}
  \label{fig:B2_A1}
\end{figure}
On the Figure \ref{fig:B2_A1} we have also shown the set of points $\omega(\mu+\rho),\; \omega\in W$ of fundamental representation of $B_2$ with the corresponding determinants of Weyl reflections $\epsilon(\omega)$. 
Now we have to factorise the Weyl group $W$ by $W_{\bot}=\left\{\omega_1\right\}$. We get the following set of anomalous points $\omega(\mu+\rho)-\rho,\; \omega\in W_{\bot}\backslash W$:
\begin{figure}[ph]
  \noindent\centering{
    \includegraphics[width=90mm]{B2_A1_2.pdf}
  }
  \caption{Anomalous points and the corresponding $\mathfrak{a}_{\bot}=A_1$-modules}
  \label{fig:B2_A1_2}
\end{figure}
We have also depicted the corresponding $\mathfrak{a}_{\bot}=A_1$-modules $L^{\pi_{\mathfrak{a}_{\bot}}(\omega(\mu+\rho))-\rho_{\mathfrak{a}_{\bot}}}_{\mathfrak{a}_{\bot}}$.
Then we project these points and dimensions of modules onto the root space of subalgebra $\mathfrak{a}=A_1$ and get the following anomalous points in fundamental weights basis with corresponding multiplicities:
\begin{equation}
  \label{eq:25}
  (1,2),\; (0,-3),\; (-4,3),\; (-5,-2)
\end{equation}
For the function $s(\gamma)$ and the set $\Gamma$ from the definition (\ref{phi-d},\ref{fan-defined}) we have
\begin{equation}
  \label{eq:22}
  (1,2),\; (2,-1)
\end{equation}
Here the second component denotes the value of $s(\gamma)$.
\begin{comment}
  It seems that here $\Gamma$ should include only two points. Is it true?

  Think of the pictures more.
\end{comment}
Anomalous branching coefficient $k^{(1,0)}_{1}=2$, then for anomalous branching coefficient $k^{(1,0)}_{0}$ the formula (\ref{recurrent relation}) gives us
\begin{equation}
  \label{eq:23}
  k^{(1,0)}_{0}=-1\cdot k^{(1,0)}_2 +2\cdot k^{(1,0)}_1 - 3\cdot \delta_{0,0} = 1
\end{equation}
So we have computed the branching coefficients.
\subsubsection{Embedding of $B_2$ into $B_4$}
\label{sec:someth-high-dimens}
Consider the the regular embedding of the subalgebra $B_2$ into the algebra $B_4$.
We calculate the branching coefficients for the fundamental representation of $B_4$.
The corresponding Dynkin diagrams are in the Figure \ref{fig:dynkin}.
\begin{figure}[ph]
  \centering
  \includegraphics[width=60mm]{B4_B2_2A1.pdf}
  \caption{Dynkin diagrams}
  \label{fig:dynkin}
\end{figure}
